---
title: "Assignment #2: Getting some data"
author: "Team 3: Marianna Garcia, Yukong Li, Jiaqi Zhu, and Mario Rodriguez-Heredia"
date: "May 7th, 2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
require(sjmisc)
```

## Loading the data

Find data from at least two sources that the members of your team agree could be interesting. The data you choose should take some wrangling to get into a flat table format. Load the data into R and merge them into a single data frame (or tibble  in the language of tidyverse).

**Answer:**
We use the Global Terorism Data set (https://www.kaggle.com/START-UMD/gtd) to create a panel data set where we add different
variables to the country-year level (e.g., number of terrorist attacks, number of causalties). Then, we joined this data with precipitation and temperature countries from the World Bank's Climate Change Knowledge Portal (https://climateknowledgeportal.worldbank.org/download-data). Finally, we include economic variables for each country using the package "pwt9" with the data of the Penn World Table V9.1.

### Step by Step:

**1.** We load terrorism microdata containing all terrorist attacks recorded between 1970 and 2017.

```{r a, warning=FALSE, message=FALSE}
terrorism <- read_csv('./data_terr/globalterrorismdb_0718dist.csv')
```
**2.** We clean some data and aggregate the individual records to data by country and year (whole code chunk in the markdown file included with this PDF). The new tibble `terr_x_ctry` contains variables such as number of attacks, fatalities, number of individuals wounded, attacks against the government or religious groups, data on weapons used, among others.

```{r b, include=FALSE}

# Reclassify location of some attacks to match with temperature and econ data
terrorism <- terrorism %>% 
  mutate( country_txt = ifelse(country_txt == "East Germany (GDR)", "Germany", country_txt),
          country_txt = ifelse(country_txt == "West Germany (FRG)", "Germany", country_txt),
          country_txt = ifelse(country_txt == "South Yemen", "Yemen", country_txt),
          country_txt = ifelse(country_txt == "Myanmar", "Myanmar (Burma)", country_txt),
          country_txt = ifelse(country_txt == "Republic of the Congo", "Congo (Republic of the)", country_txt),
          country_txt = ifelse(country_txt == "People's Republic of the Congo", "Congo (Republic of the)", country_txt),
          country_txt = ifelse(country_txt == "Democratic Republic of the Congo", "Congo (Republic of the)", country_txt),
          country_txt = ifelse(country_txt == "Zaire", "Congo (Democratic Republic of the)", country_txt),
          country_txt = ifelse(country_txt == "North Yemen", "Yemen", country_txt),
          country_txt = ifelse(country_txt == "South Korea", "Korea" , country_txt),
          country_txt = ifelse(country_txt == "Rhodesia", "Zimbabwe", country_txt),
          country_txt = ifelse(country_txt == "Vatican City", "Italy", country_txt),
          country_txt = ifelse(country_txt == "Ivory Coast", "Cote d'Ivoire", country_txt),
          country_txt = ifelse(country_txt == "Montenegro", "Republic of Montenegro", country_txt),
          country_txt = ifelse(country_txt == "Serbia", "Republic of Serbia", country_txt),
          country_txt = ifelse(country_txt == "Slovak Republic", "Slovakia", country_txt),
          country_txt = ifelse(country_txt == "South Vietnam", "Vietnam", country_txt),
          country_txt = ifelse(country_txt == "Western Sahara", "Morocco", country_txt),
          country_txt = ifelse(country_txt == "Bosnia-Herzegovina", "Bosnia and Herzegovina", country_txt),
          )

# Create indicator variables and group attacks by country-year
terrorism <- terrorism %>% 
  mutate(counter = 1,
         atktrg_gov = ifelse(targtype1_txt == "Government (General)",1,0 ),
         atktrg_diplom = ifelse(targtype1_txt == "Government (Diplomatic)",1,0 ),
         atktrg_relig = ifelse(targtype1_txt == "Religious Figures/Institutions",1,0 ),
         atktrg_air = ifelse(targtype1_txt == "Airports & Aircraft",1,0 ),
         atktrg_military = ifelse(targtype1_txt == "Military",1,0 ),
         atktrg_buss = ifelse(targtype1_txt == "Business",1,0 ),
         atktrg_priv = ifelse(targtype1_txt == "Private Citizens & Property",1,0 ),
         with_explosives = ifelse(weaptype1_txt == "Explosives",1,0 ),
         with_sabotage = ifelse(weaptype1_txt == "Sabotage Equipment",1,0 ),
         with_firearm = ifelse(weaptype1_txt == "Firearms",1,0 ))

# summarize main variables by country-year
terr_x_ctry <-  terrorism %>% group_by(iyear, country_txt) %>% 
  summarize( number_of_attacks = sum(counter), # Total number of terrorist attacks
             eventsLT1hr = sum(extended, na.rm = T), # terrorist events that last more than 1 hr
             success = sum(success, na.rm = T), # successful terrorist attacks
             suicide_attack = sum(suicide, na.rm = T), # suicide attacks
             confirmed_fatalities = sum(nkill, na.rm = T), # confirmed fatalities
             confirmed_us_fatal = sum(nkillus, na.rm = T), # confirmed US-Citizen fatalities
             nkillter = sum(nkillter, na.rm = T), # confimed terrorist fatalities
             nwound = sum(nwound, na.rm=T), # number of wounded 
             nwoundus = sum(nwoundus, na.rm = T), # number of US-Citizens wounded
             nwoundte = sum(nwoundte, na.rm = T), # number of terrorist wounded
             involved_property = sum(property==1, na.rm = T), # number of attacks on property
             ransom = sum(ransom==1, na.rm = T), # attacks were ransom was demanded
             atktrg_gov = sum(atktrg_gov), # number of attacks on government (general)
             atktrg_diplom = sum(atktrg_diplom), # number of attacks on government (diplomatic)
             atktrg_relig = sum(atktrg_relig), # attack on religious figures/institutions
             atktrg_air = sum(atktrg_air), # number of attacks on aircraft or airports
             atktrg_military = sum(atktrg_military), # number of attacks on military
             atktrg_buss = sum(atktrg_buss), # ... on business
             atktrg_priv = sum(atktrg_priv), # ... on private property and citizens
             with_explosives = sum(with_explosives), # attacks where explosives were used
             with_sabotage = sum(with_sabotage), # attacks where sabotage equipment was used
             with_firearm = sum(with_firearm) # attacks with firearms
             ) %>% 
  rename( country = country_txt, year = iyear)
```

```{r c, echo=TRUE}
terr_x_ctry
```

***3.*** We join the aggregate data with an empty panel because countries with no events in specific years will not have observations in those years. Then, we substitute the NAs values in the full panel data set with zeros since an NA means there were no events recorded that year (whole code chunk in the markdown).
```{r panel_data}
# create empty panel data 
panel <- tibble(year = rep(unique(terr_x_ctry$year), times = 195),
                country = rep(unique(terr_x_ctry$country), each = 47))

# join the empty panel with the aggregated terrorism data
panel <- panel %>% 
  left_join(terr_x_ctry, by = c('year' , 'country'))
```

````{r replace_NA, include = FALSE}
# Replace NA for 0, since NA means that there were no attacks that month/year in
# that country
panel <- panel %>% 
  mutate( number_of_attacks = ifelse(is.na(number_of_attacks), 0, number_of_attacks),
          eventsLT1hr = ifelse(is.na(eventsLT1hr), 0, eventsLT1hr),
          success = ifelse(is.na(success), 0, success),
          suicide_attack = ifelse(is.na(suicide_attack), 0, suicide_attack),
          confirmed_fatalities = ifelse(is.na(confirmed_fatalities), 0, confirmed_fatalities),
          confirmed_us_fatal = ifelse(is.na(confirmed_us_fatal), 0, confirmed_us_fatal),
          nkillter = ifelse(is.na(nkillter), 0, nkillter),
          nwound = ifelse(is.na(nwound), 0, nwound),
          nwoundus = ifelse(is.na(nwoundus), 0, nwoundus),
          nwoundte = ifelse(is.na(nwoundte), 0, nwoundte),
          involved_property = ifelse(is.na(involved_property), 0, involved_property),
          ransom = ifelse(is.na(ransom), 0, ransom),
          atktrg_gov = ifelse(is.na(atktrg_gov), 0, atktrg_gov),
          atktrg_diplom = ifelse(is.na(atktrg_diplom), 0, atktrg_diplom),
          atktrg_relig = ifelse(is.na(atktrg_relig), 0, atktrg_relig),
          atktrg_air = ifelse(is.na(atktrg_air), 0, atktrg_air),
          atktrg_military = ifelse(is.na(atktrg_military), 0, atktrg_military),
          atktrg_buss = ifelse(is.na(atktrg_buss), 0, atktrg_buss),
          atktrg_priv = ifelse(is.na(atktrg_priv), 0, atktrg_priv),
          with_explosives = ifelse(is.na(with_explosives), 0, with_explosives),
          with_sabotage = ifelse(is.na(with_sabotage), 0, with_sabotage),
          with_firearm = ifelse(is.na(with_firearm), 0, with_firearm))


#add_region
regions_by_country <- terrorism %>% filter(!is.na(region_txt)) %>% 
  group_by(country_txt) %>% 
  summarize(region = first(region_txt)) %>% 
  rename(country = country_txt)

panel <-   panel %>% left_join(regions_by_country, by = "country" )

# arrange columns
panel <- move_columns(panel,  region, .after = country )
```
***4.***  We load the temperature and precipitation data and aggregate it at the country-year level. Then, we joined these sets with the terrorism tibble.

```{r join_pr_rn, echo =TRUE, message=FALSE, warning=FALSE}
#load precipitation and temperature data
precipitation <- read_csv('./precipitation_1901_2020.csv') %>% 
  rename( country = Country, year = Year ) 

temperature <- read_csv('./temperature_1901_2020.csv') %>% 
  rename( country = Country, year = Year )

# summarize data by year and country
precipitation_yr <- precipitation %>%  group_by(country, year) %>% 
  summarize(av_rainfall_mm = mean(`Rainfall - (MM)`, na.rm=T),
            max_rainfall_mm = max(`Rainfall - (MM)`, na.rm=T),
            min_rainfall_mm = min(`Rainfall - (MM)`, na.rm=T),
            sd_rainfall_mm = sd(`Rainfall - (MM)`, na.rm=T),
            ISO3 = first(ISO3))

temperature_yr <- temperature %>%  group_by(country, year) %>% 
  summarize(av_temp_c = mean(`Temperature - (Celsius)`, na.rm=T),
            max_temp_c = max(`Temperature - (Celsius)`, na.rm=T),
            min_temp_c = min(`Temperature - (Celsius)`, na.rm=T),
            sd_temp_c = sd(`Temperature - (Celsius)`, na.rm=T))

# merge
panel_df <- panel %>% left_join(precipitation_yr, by = c('year',  'country')) %>% 
  left_join(temperature_yr, by = c('year', 'country'))

```

```{r move_columns, include=FALSE}
panel_df <- move_columns(panel_df,  ISO3, .after = country )
panel_df <- move_columns(panel_df,  av_rainfall_mm, .after = number_of_attacks )
panel_df <- move_columns(panel_df,  av_temp_c, .after = av_rainfall_mm)
panel_df <- tibble(panel_df)


#  remaining cleaning issues
check <- panel_df %>%  filter_all(any_vars(is.na(.)))
unique(check$country)
na_countries <- unique(check$country)
sum(check$number_of_attacks) 
```
***5.*** We merge with the PWT9.1, arrange the columns, and do some last cleaning (whole code chunk in markdown file).

```{r PWT, message=FALSE, warning=FALSE}
# Data for Pen Wold Table
require(pwt9)
pwt_lite <- pwt9.1 %>% select(country, ISO3 = isocode, year, rgdpe, 
                              pop, emp, hc, labsh, xr, pl_con, csh_c,
                              csh_g, csh_i, csh_x, csh_m) %>% 
  filter(year > 1969)

# merge
panel_df_vf <- panel_df %>% left_join(pwt_lite, by = c('year', 'country')) 
```

```{r cleaning_final, include=FALSE}
panel_df_vf <- tibble(
  move_columns(panel_df_vf, c('rgdpe', 'pop') , .after = av_temp_c))


# eliminate countries were there is some missing data of precipitation
panel_df_vf <- panel_df_vf %>% filter( !(country %in% na_countries) )

panel_df_vf <-  panel_df_vf %>%  select(-c("ISO3.y")) %>% arrange(year) %>% 
  arrange(desc(country))
```

```{r show, echo=TRUE}
panel_df_vf
```


## Long to wide

What is the unit of analysis or aggregation?  Can these data be switched between long and wide formats?  If so, do this and then identify the unit of analysis in the new dataset.

**Answer:** We have a panel with years as the time variable and country as the cross-sectional, then our unit of observation is a country in a given year. Currently, our data is in long format, and we can switch it to a wide format, as shown below. In the wide-format, the unit of analysis is the country. However, since we have more than 40 variables, the number of columns increases significantly, and the data may be harder to visualize.

```{r wide}

# select time invariant variables
time_invariant_var <- names(panel_df_vf)
time_invariant_var <- time_invariant_var[!time_invariant_var
                                         %in% c("year", "country", "ISO3.x",
                                                "region" )]

# switch to wide
wide_df <- panel_df_vf %>%  pivot_wider(
  names_from = c("year"),
  values_from = all_of(time_invariant_var) ) %>% arrange(country)

wide_df[,1:30] # show first 30 columns/variables only
```

